---
title: "Practical Machine Learning-Stream 005 Course Project Write Up"
author: "Jenny"
date: "22 January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset)."

Data

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. The information has been generously provided for use use in this cousera course by the authors, Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. They have allowed the use of their paper "Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

Both were downloaded for use in this assignment on September 3, 2015 at 6:50am GMT + 8hours.


```{r load_data}
library(ggplot2)
library(caret)

library(randomForest)

library(e1071)
library(gbm)

library(doParallel)
library(survival)
library(splines)
library(plyr)

training <- read.csv("C:/Users/HP/Documents/R/course8/pml-training.csv", na.strings=c("#DIV/0!"), row.names = 1)
testing <- read.csv("C:/Users/HP/Documents/R/course8/pml-testing.csv", na.strings=c("#DIV/0!"), row.names = 1)
training <- training[, 6:dim(training)[2]]

treshold <- dim(training)[1] * 0.95
#Remove columns with more than 95% of NA or "" values
goodColumns <- !apply(training, 2, function(x) sum(is.na(x)) > treshold  || sum(x=="") > treshold)

training <- training[, goodColumns]

badColumns <- nearZeroVar(training, saveMetrics = TRUE)

training <- training[, badColumns$nzv==FALSE]

training$classe = factor(training$classe)

#Partition rows into training and crossvalidation
inTrain <- createDataPartition(training$classe, p = 0.6)[[1]]
crossv <- training[-inTrain,]
training <- training[ inTrain,]
inTrain <- createDataPartition(crossv$classe, p = 0.75)[[1]]
crossv_test <- crossv[ -inTrain,]
crossv <- crossv[inTrain,]


testing <- testing[, 6:dim(testing)[2]]
testing <- testing[, goodColumns]
testing$classe <- NA
testing <- testing[, badColumns$nzv==FALSE]
#Train 3 different models
mod1 <- train(classe ~ ., data=training, method="rf")
#mod2 <- train(classe ~ ., data=training, method="gbm")
#mod3 <- train(classe ~ ., data=training, method="lda")

pred1 <- predict(mod1, crossv)
#pred2 <- predict(mod2, crossv)
#pred3 <- predict(mod3, crossv)
#show confusion matrices
confusionMatrix(pred1, crossv$classe)
